\section{Safety and Risks}

We are committed to ensuring that GLM-4 operates as a safe, responsible, and unbiased model. 
In addition to addressing common ethical and fairness concerns, we carefully assess and mitigate potential harms that the model may pose to users in real-world scenarios.

\input{tables/safetytable}

\vpara{Risk Mitigation.}
We carefully clean data in the pre-training stage by removing text containing sensitive keywords and web pages from a pre-defined blacklist.
In the alignment phase, we evaluate each training sample for safety and remove any that pose potential risks.
Harmlessness is also an important criteria for preference alignment when comparing multiple model outputs.

We have a red team that constantly challenges the model with tricky questions that tend to cause unsafe answers.
We collect all harmful question-answer pairs from GLM-4 and improve them with human annotations for further model alignment.

\newcommand{\safetyfont}[1]{{\emph{#1}}}

\vpara{Safety Evaluation.}
We evaluate the GLM-4 model on the SafetyBench \cite{zhang2023safetybench} dataset, which assesses the capability of each model from 7 dimensions: 
\safetyfont{Ethics and Morality} (unethical behaviors), 
\safetyfont{Illegal Activities} (basic knowledge of law), 
\safetyfont{Mental Health} (adverse impacts on mental health), 
\safetyfont{Offensiveness} (offensive behaviors), 
\safetyfont{Physical Health} (dangerous behaviors that can cause physical harms), 
\safetyfont{Privacy and Property} (privacy breach or property loss), 
\safetyfont{Unfairness and Bias}.
We evaluate different models on the Chinese subset of SafetyBench, which is created by removing highly sensitive questions that tend to be blocked, to mitigate interference from different API safety policies.

Table \ref{tb:safety} shows the safety results of GLM-4 and SOTA models.
On most dimensions GLM-4 (0520) shows competitive safety performance, and overall it achieves comparable performance with Claude 3 Opus.
GLM-4 slightly falls behind the GPT-4 family of models, especially on the Physical Health dimension, which demands robust common sense knowledge about the physical world to avoid potential risks.
More efforts have been put into this direction to develop a more capable and safe GLM model.
